---
title: "Lab 1"
author: "Óscar Agüero Rodriguez"
date: "30/8/2020"
output: 
  github_document:
    toc: TRUE
    toc_depth: 2
---

Paquetes a utilizar

```{r}
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(sf))
suppressPackageStartupMessages(library(mapview))
suppressPackageStartupMessages(library(abind))
suppressPackageStartupMessages(library(stars))
suppressPackageStartupMessages(library(RSQLite))
suppressPackageStartupMessages(library(spacetime))
suppressPackageStartupMessages(library(xts))
suppressPackageStartupMessages(library(starsdata))
suppressPackageStartupMessages(library(spData))
suppressPackageStartupMessages(library(spDataLarge))
suppressPackageStartupMessages(library(units))
suppressPackageStartupMessages(library(cubelyr))
suppressPackageStartupMessages(library(tmap))
```

# 1.3 Exercises

## 1.3.1 

Read the shapefile storms_xyz_feature from the shape directory in the sf package

```{r}
tst = st_read(system.file("shape/storms_xyz_feature.shp", package="sf"), quiet = TRUE)
class(st_geometry(tst))
class(st_geometry(tst)[[1]])
```


## 1.3.2 

Copy this file to another directory on your computer, and read it from there (note: a shapefile consists of more than one file!)

```{r}
setwd("~/Estadística/Spacial Stats/Storms_xyz")
tst1 <- st_read("storms_xyz_feature.shp")
```

## 1.3.3 

How many features does this dataset contain?

```{r}
tst1
```

Tiene 71 features

## 1.3.4 

Plot the dataset, with axes = TRUE (hint: before plotting, pipe through st_zm to drop Z and M coordinates; more about this in chapter 3).

```{r}
system.file("shape/storms_xyz_feature.shp", package="sf") %>%
  read_sf() %>%
  st_zm() %>% 
  plot(graticule = TRUE, axes = TRUE)

```

## 1.3.5 

Before plotting, pipe the dataset through st_set_crs(4326). What is different in the plot obtained?

```{r}
system.file("shape/storms_xyz_feature.shp", package="sf") %>%
  read_sf() %>%
  st_zm() %>% 
  st_set_crs(4326) %>% 
  plot(graticule = TRUE, axes = TRUE)
```

La diferencia observada esta principalmente en las coordenadas del plano, en el segunndo plano se pueden ver la dirección de los grados en los margenes.

# 2.6 Exercices

## 2.6.1 

Convert $(x , y)$ point (10,2), (-10,-2), (10,2) and (0,10) to polar coordinates

Coordenada (10,2)

```{r}
x1 <- 10; y1 <- 2
c1 <- sqrt(x1^2 + y1^2)
p1 <- atan(y1/x1)
copor1 <- cbind(c1, p1);copor1

```

Coordenada (-10,-2)

```{r}
x2 <- -10; y2 <- -2
c2 <- sqrt(x2^2 + y2^2)
p2 <- atan(y2/x2)
copor2 <- cbind(c2, p2);copor2
```

Coordenada (10,-2)

```{r}
x3 <- 10; y3 <- -2
c3 <- sqrt(x3^2 + y3^2)
p3 <- atan(y3/x3)
copor3 <- cbind(c3, p3);copor3
```

Coordenada (0,10)

```{r}
x4 <- 0; y4 <- 10
c4 <- sqrt(x4^2 + y4^2)
p4 <- atan(y4/x4)
copor4 <- cbind(c4, p4);copor4

```

## 2.6.2

Convert the polar $(r,\phi)$ points (10, 45°), (0,100°) and (5,359°) to cartesian coordinates

```{r}
r1 <- 10;phi1 <- 45
r2 <- 0; phi2 <- 100
r3 <- 5; phi3 <- 359

x_1 <- r1*cos(phi1)
y_1 <- r1*sin(phi1)

x_2 <- r2*cos(phi2)
y_2 <- r2*sin(phi2)

x_3 <- r3*cos(phi3)
y_3 <- r3*sin(phi3)

cc1 <- round(cbind(x_1,y_1),1)
cc2 <- round(cbind(x_2,y_2),1)
cc3 <- round(cbind(x_3,y_3),1)

resumen <- cbind(cc1, cc2, cc3)
resumen

```

## 2.6.3

Assuming the Earth is a sphere with a radius of 6371 km, compute for $(\lambda, \phi)$ points the great circle distance between (10,10) and (11,10), between (10,80) and (11,80), between (10,10) and (11,10) and between (10,80) and (11,81) (units: degree): What the distance units?

Para dos punto $p_1 = (\lambda_1, \phi_1)$ y $p_2 = (\lambda_2, \phi_2)$, donde $p_1$ y $p_2$ deben ser calculados en radianes, se debe convertir primero los grados a Radianes.

Eso se realiza entendiendo que:

$$1° = \frac{\pi}{180}$$

Conversión

```{r}
lambda1 <- 10*(pi/180)
theta1 <- 10*(pi/180)
lambda2 <- 11*(pi/180)
theta2 <- 10*(pi/180)

lambda3 <- 10*(pi/180)
theta3 <- 80*(pi/180)
lambda4 <- 11*(pi/180)
theta4 <- 80*(pi/180)

lambda5 <- 10*(pi/180)
theta5 <- 10*(pi/180)
lambda6 <- 10*(pi/180)
theta6 <- 11*(pi/180)

lambda7 <- 10*(pi/180)
theta7 <- 80*(pi/180)
lambda8 <- 10*(pi/180)
theta8 <- 81*(pi/180)

r <- 6371
```

Para obtener la distancia además, se debe aplicar la siguiente formulas en esta secuencia

$$\theta_{12} = arccos(sin(\theta_1)*sin(\theta_2)+cos(\theta_1)*cos(\theta_2)cos(|\lambda_1 - \lambda_2|))$$

$$S_{12} = r*\theta_{12}$$

Distancia (1,2) entre arco (10,10) y (11,10)

```{r}
theta12 <- acos(sin(theta1)*sin(theta2)+cos(theta1)*cos(theta2)*cos(abs(lambda1-lambda2)))
s12 <- r*theta12
s_12g <- (s12*180)/pi 

s12;s_12g

```

Distancia (3,4) entre arco (10,80) y (11,80)

```{r}
theta34 <- acos(sin(theta3)*sin(theta4)+cos(theta3)*cos(theta4)*cos(abs(lambda3-lambda4)))
s34 <- r*theta34
s_34g <- (s34*180)/pi 

s34;s_34g

```

Distancia (5,6) entre arco (10,10) y (10,11)

```{r}
theta56 <- acos(sin(theta5)*sin(theta6)+cos(theta5)*cos(theta6)*cos(abs(lambda5-lambda6)))
s56 <- r*theta56
s_56g <- (s56*180)/pi 

s56;s_56g
```

Distancia (7,8) entre arco (10,80) y (10,81)

```{r}
theta78 <- acos(sin(theta7)*sin(theta8)+cos(theta7)*cos(theta8)*cos(abs(lambda7-lambda8)))
s78 <- r*theta78
s_78g <- (s78*180)/pi 

s78;s_78g
```

# 4.4 Exercises 

## 4.4.1

NDVI, normalized differenced vegetation index, is computed as (NIR-R)/(NIR+R), with NIR the near infrared and R the red band. Read the L7_ETMs.tif file into object x, and distribute the band dimensions over attributes by split(x,"band"). Then, compute NDVI by using a expression that uses the NIR (band 4) and R (band 3) attributes directly.

```{r}
tif = system.file("tif/L7_ETMs.tif", package = "stars")
x = read_stars(tif)
split(x,"band")

```

```{r}
ndvi = function(x) (x[4]-x[3])/(x[4]+x[3])
st_apply(x, c("x", "y"), ndvi)
```
## 4.4.2

Compute NDVI for the S2 image, using st_apply and an function ndvi = function(x) (x[4]-x[3])/(x[4]+x[3]). Plot the result, and write the result to a GeoTIFF. Explian the difference in runtime between plotting and writing.

```{r}
ndvi = function(x) (x[4]-x[3])/(x[4]+x[3])
y <- st_apply(x, c("x", "y"), ndvi)
y
```

Plotting and writting as GeoTIFF

```{r}
tf = tempfile(fileext=".tif")
write_stars(y, tf)
st_drivers("raster")
plot(y)
```

## 4.4.3

Use st_transform to transform the stars object read from L7_ETMs.tif to EPSG 4326. Print the object. Is this a regular grid? Plot the first band using arguments axes=TRUE and border=NA, and explain why this takes such a long time.

```{r}
tif = system.file("tif/L7_ETMs.tif", package = "stars")
x = read_stars(tif)
new = st_crs(4326)
y <- st_transform(x, new)
y

```

```{r}

plot(y, axes = TRUE, border =NA)

```

Toma mucho tiempo porque los datos de tif/L7_ETMs.tif son tipo raster y al aplicarle una transformación le estamos cambiando a otro sistema de coordenadas lo que genera perdida de datos.

## 4.4.4

Use st_warp to warp the L7_ETMs.tif object to EPSG 4326, and plot the resulting object with axes=TRUE. Why is the plot created much faster than after st_transform?

```{r}
tif <- system.file("tif/L7_ETMs.tif", package = "stars")
x <- read_stars(tif)
new_crs <- st_crs(4326)
y <- st_warp(x,crs = new_crs)
y
```

```{r}
plot(y, axes = TRUE)
```

Se tiene mejor eficiencia comparado a utilizar st_transform debido a que no se le modifica el tipo de referencia de coordenada.


# 6.6 Exercises

## 6.6.1

Add a variable to the nc dataset by nc$State = "North Carolina". Which value should you attach to this variable for the attribute-geometry relationship (agr)?

```{r}
nc <- system.file("gpkg/nc.gpkg", package="sf") %>%
    read_sf() %>%
    st_transform(32119)

nc <- nc %>% 
  mutate(State = "North Carolina")

```

Se le debe ligar a una geometría de centroide, dado que no se cuenta con una especifica para North Carolina, para esto se puede utilizar st_agr


## 6.6.2

Create a new sf object from the geometry obtained by st_union(nc), and assign "North Carolina" to the variable State. Which agr can you now assign to this attribute variable?

```{r}
nc <- system.file("gpkg/nc.gpkg", package="sf") %>%
    read_sf() %>%
    st_transform(32119)
nc <- nc %>% 
  mutate(State = "North Carolina")
st_union(nc)
```

```{r}
nc1 <- nc %>% select(BIR74, SID74, NAME, State) %>%
    st_set_agr(c(BIR74 = "aggregate", SID74 = "aggregate", NAME = "identity", State = "identity"))
nc1
```

Dado que todos pertenencen al estado de North Carolina, solo se requiere aplicar el valor "Identity" para asignar un valor unico a la geometria.

## 6.6.3

Use st_area to add a variable with name area to nc. Compare the area and AREA variables in the nc dataset. What are the units of AREA? Are the two linearly related? If there are discrepancies, what could be the cause?

```{r}
nc$area <- st_area(nc)
nc1 <- head(cbind(nc$AREA, nc$area))
nc2 <- tail(cbind(nc$AREA, nc$area))
colnames(nc1) <- c("AREA", "area"); nc1
```

```{r}
colnames(nc2) <- c("AREA", "area"); nc2

```

Las unidades de AREA son muy similares a los de area, sin embargo, los de AREA parecen que salieran de la división de area entre 10 millones.

Dado que ambas parecen tener los mismos valores, solo que una diferente escala puede decirse que si hay linearidad. Sin embargo, cabe resaltar que la similutud de los datos se da hasta el 3 decimal, ya que de ahí en adelante si pareciera existir una discrepancia en los datos.

## 6.6.4

Is the area variable intensive or extensive? Is its agr equal to constant, identity or aggregate?

En este caso se tiene un area por lo que se tiene que utilizar extensive. Además, al ser una cantidad se debe utilizar el aggregate al usar agr.


## 6.6.5

Find the name of the county that contains POINT(-78.34046 35.017)

```{r}
nc <- system.file("gpkg/nc.gpkg", package="sf") %>%
    read_sf() %>%
    st_transform(4326)

punto <- st_point( c( -78.34046, 35.017 ) )
res <- st_contains(nc, punto)
res <- summary(res) 
res_1 <- as.numeric( res[,1])
pos <- which( res_1==1 )
nc$NAME[pos]

```

El punto identificado es Sampson.

## 6.6.6

Find the names of all counties with boundaries that touch county Sampson

```{r}
sampson_geo <- nc$geom[pos]

counties_2 <- st_intersection(nc, sampson_geo)
counties_2 <- counties_2[ counties_2$NAME!='Sampson',    ]; counties_2$NAME
```

## 6.6.7

List the names of all counties that are less than 50 KM away from county Sampson

```{r}
counties_3 <- st_is_within_distance(nc, sampson_geo, dist = 50000)
counties_3 <- summary(counties_3) 
counties_3 <- as.numeric(counties_3[,1])
pos <- which( counties_3==1 )
nc$NAME[pos]
```

# 8.7 Exercises

```{r}
africa = world %>% 
   filter(continent == "Africa", !is.na(iso_a2)) %>% 
   left_join(worldbank_df, by = "iso_a2") %>% 
   dplyr::select(name, subregion, gdpPercap, HDI, pop_growth) %>% 
   st_transform("+proj=aea +lat_1=20 +lat_2=-23 +lat_0=0 +lon_0=25")
```


```{r}
zion = st_read((system.file("vector/zion.gpkg", package = "spDataLarge")))

data(nlcd, package = "spDataLarge")
```


## 8.7.1

Create a map showing the geographic distribution of the Human Development Index (HDI) across Africa with base graphics (hint: use plot()) and tmap packages (hint: use tm_shape(africa) + ...).

Usando plot()

```{r}
plot(africa[4])
```

Usando tm_shape

```{r}
tm_shape(africa)+ tm_polygons(col = "HDI")
```

Ventaja 1: con tm_shape la programación es similar a ggplot2 lo cual permite tener claridad en lo que se quiere programar de manera lógica.

Ventaja 2: Al utilizar plot, no fue directo que se logra tener el mapa deseado, primero tuve que probar hasta encontrar (esto no lo reflejo en el código), con tm_shape se puede hacer directo lo que se busca al gráficar.

Otras formas de graficar pueden ser con ggplot2, cartogram y mapview todas son librerias de R, por lo cual se pueden utilizar sin tener que agregar costo a un trabajo o una investigación.

Bonus

Usando cartogram

```{r}
library(cartogram)
africa_carto = cartogram_ncont(africa, "HDI")
tm_shape(africa_carto) + tm_polygons(col = "HDI")

```

Usando ggplot2

```{r}
g1 <- ggplot() + geom_sf(data = africa, aes(fill = HDI))
g1
```

## 8.7.2

Extend the tmap created for the previous exercise so the legend has three bins: “High” (HDI above 0.7), “Medium” (HDI between 0.55 and 0.7) and “Low” (HDI below 0.55).

```{r}

africa = africa %>% mutate(rango_HDI=case_when(HDI<0.55~'bajo',
                                               HDI>=0.55 & HDI<0.7  ~'medio',
                                               HDI>0.7 ~'alto',
                                               TRUE~'nd'))

map1 <- tm_shape(africa)+ tm_fill(col = "rango_HDI") + tm_borders(col = "black") + tm_layout(title = "Africa por HDI") + tm_layout(scale = 2) + tm_layout(bg.color = "lightblue") + tm_compass(type = "8star", position = c("right", "top"))

map1
```

## 8.7.3

Represent africa’s subregions on the map. Change the default color palette and legend title. Next, combine this map and the map created in the previous exercise into a single plot.


```{r}
map2 <- tm_shape(africa)+ tm_fill(col = "subregion", palette = "Set2") + tm_layout(title = "Africa por subregiones")
map2
```

```{r}
tmap_arrange(map1, map2)
```

## 8.7.4

Create a land cover map of the Zion National Park.





