---
title: "Lab 1"
author: "Óscar Agüero Rodriguez"
date: "30/8/2020"
output: 
  github_document:
    toc: TRUE
    toc_depth: 2
---

Paquetes a utilizar

```{r}
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(sf))
suppressPackageStartupMessages(library(mapview))
suppressPackageStartupMessages(library(abind))
suppressPackageStartupMessages(library(stars))
suppressPackageStartupMessages(library(RSQLite))
suppressPackageStartupMessages(library(spacetime))
suppressPackageStartupMessages(library(xts))
suppressPackageStartupMessages(library(starsdata))
suppressPackageStartupMessages(library(spData))
suppressPackageStartupMessages(library(spDataLarge))
suppressPackageStartupMessages(library(units))
suppressPackageStartupMessages(library(cubelyr))
suppressPackageStartupMessages(library(tmap))
```

# Lab 1.3

## 1.3.1 

Read the shapefile storms_xyz_feature from the shape directory in the sf package

```{r}
tst = st_read(system.file("shape/storms_xyz_feature.shp", package="sf"), quiet = TRUE)
class(st_geometry(tst))
class(st_geometry(tst)[[1]])
```


## 1.3.2 

Copy this file to another directory on your computer, and read it from there (note: a shapefile consists of more than one file!)

```{r}
setwd("~/Estadística/Spacial Stats/Storms_xyz")
tst1 <- st_read("storms_xyz_feature.shp")
```

## 1.3.3 

How many features does this dataset contain?

```{r}
tst1
```

Tiene 71 features

## 1.3.4 

Plot the dataset, with axes = TRUE (hint: before plotting, pipe through st_zm to drop Z and M coordinates; more about this in chapter 3).

```{r}
system.file("shape/storms_xyz_feature.shp", package="sf") %>%
  read_sf() %>%
  st_zm() %>% 
  plot(graticule = TRUE, axes = TRUE)

```

## 1.3.5 

Before plotting, pipe the dataset through st_set_crs(4326). What is different in the plot obtained?

```{r}
system.file("shape/storms_xyz_feature.shp", package="sf") %>%
  read_sf() %>%
  st_zm() %>% 
  st_set_crs(4326) %>% 
  plot(graticule = TRUE, axes = TRUE)
```

La diferencia observada esta principalmente en las coordenadas del plano, en el segunndo plano se pueden ver la dirección de los grados en los margenes.

# 2.6 Exercices

## 2.6.1 

Convert $(x , y)$ point (10,2), (-10,-2), (10,2) and (0,10) to polar coordinates

Coordenada (10,2)

```{r}
x1 <- 10; y1 <- 2
c1 <- sqrt(x1^2 + y1^2)
p1 <- atan(y1/x1)
copor1 <- cbind(c1, p1);copor1

```

Coordenada (-10,-2)

```{r}
x2 <- -10; y2 <- -2
c2 <- sqrt(x2^2 + y2^2)
p2 <- atan(y2/x2)
copor2 <- cbind(c2, p2);copor2
```

Coordenada (10,-2)

```{r}
x3 <- 10; y3 <- -2
c3 <- sqrt(x3^2 + y3^2)
p3 <- atan(y3/x3)
copor3 <- cbind(c3, p3);copor3
```

Coordenada (0,10)

```{r}
x4 <- 0; y4 <- 10
c4 <- sqrt(x4^2 + y4^2)
p4 <- atan(y4/x4)
copor4 <- cbind(c4, p4);copor4

```

## 2.6.2

Convert the polar $(r,\phi)$ points (10, 45°), (0,100°) and (5,359°) to cartesian coordinates






